Sorting algorithms: how are they affected by memory faults?

### 2019-12-22 paper/DataAnalysis.tex
finalização da escrita do artigo

In this section, we present our results after the execution of sorting algorithms over the input files. We analyzed only the dependent variables \textit{percentage of the largest subarray size ((\%LSS)} and \textit{percentage of unordered elements quantity (\%UEQ)}. These variables, because they are a percentage value, already were normalized (i.e., the same order of magnitude) related to dependent variable array size.

@@

\subsection{Dependente Variables Normality}

@@

Tables \ref{table-shapiro-test-lss} and \ref{table-shapiro-test-ueq} shows examples of results we get running Shapiro-Wilk test. The independent variables assume these values: \textit{probability of failure} of 0.01 and \textit{array size} of 100. The bold \textit{p-values} confirms that the data is normally distributed.

\begin{table}[H]
    \parbox{.45\linewidth}{
        \caption{Shapiro-Wilk test for \%LSS with \textit{probability of failure} of 1\% and \textit{array size} of 100.}
        \begin{center}
        \begin{tabular}{|c|c|c|}
        \hline
        \textbf{Algorithm} & \textbf{W} & \textbf{p-value} \\
        \hline
        Bubblesort & 0.854840 & 0.0007 \\
        \hline
        Insertion Sort & 0.961790 & \textbf{0.3439} \\
        \hline
        Mergesort & 0.937173 & \textbf{0.0763} \\
        \hline
        Quicksort & 0.869708 & 0.0016 \\
        \hline
        \end{tabular}
        \label{table-shapiro-test-lss}
        \end{center}
    }
    \hfill
    \parbox{.45\linewidth}{
        \caption{Shapiro-Wilk test for \%UEQ with \textit{probability of failure} of 1\% and \textit{array size} of 100.}
        \begin{center}
        \begin{tabular}{|c|c|c|}
        \hline
        \textbf{Algorithm} & \textbf{W} & \textbf{p-value} \\
        \hline
        Bubblesort & 0.934980 & \textbf{0.0666} \\
        \hline
        Insertion Sort & 0.949881 & \textbf{0.1678} \\
        \hline
        Mergesort & 0.936667 & \textbf{0.0739} \\
        \hline
        Quicksort & 0.936049 & \textbf{0.0712} \\
        \hline
        \end{tabular}
        \label{table-shapiro-test-ueq}
        \end{center}
    }
\end{table}

@@

\begin{figure}[H]
    \centering
    \frame{\includegraphics[scale=0.30]{figures/anova_prob_0_01_tam_100_Largest_subarray_size.png}}
    \textsf{\caption[Q-Q plot for \%LSS with \textit{probability of failure} of 1\% and \textit{array size} of 100.]{Q-Q plot for \%LSS with \textit{probability of failure} of 1\% and \textit{array size} of 100.\label{fig-qqplot-lss-001-100}}}
 \end{figure}

@@

>>>>

In this section, we present our results after the execution of sorting algorithms over the input files. We analyzed only the dependent variables \textit{percentage of the largest subarray size ((\%LSS)} and \textit{percentage of unordered elements quantity (\%UEQ)}. These variables, because they are a percentage value, already were normalized (i.e., the same order of magnitude) related to dependent variable \textit{array size}.

@@

\subsection{Dependent Variables Normality}

@@

Tables \ref{table-shapiro-test-lss-001-100} and \ref{table-shapiro-test-ueq-001-100} shows examples of results we got running Shapiro-Wilk test. The independent variables assume these values: \textit{probability of failure} of 0.01 and \textit{array size} of 100. The bold \textit{p-values} confirms that the data is normally distributed.

\begin{table}[H]
    \parbox{.45\linewidth}{
        \caption{Shapiro-Wilk test for \%LSS with \textit{probability of failure} of 1\% and \textit{array size} of 100.}
        \begin{center}
        \begin{tabular}{|c|c|c|}
        \hline
        \textbf{Algorithm} & \textbf{W} & \textbf{p-value} \\
        \hline
        Bubblesort & 0.854840 & 0.0007 \\
        \hline
        Insertion Sort & 0.961790 & \textbf{0.3439} \\
        \hline
        Mergesort & 0.937173 & \textbf{0.0763} \\
        \hline
        Quicksort & 0.869708 & 0.0016 \\
        \hline
        \end{tabular}
        \label{table-shapiro-test-lss-001-100}
        \end{center}
    }
    \hfill
    \parbox{.45\linewidth}{
        \caption{Shapiro-Wilk test for \%UEQ with \textit{probability of failure} of 1\% and \textit{array size} of 100.}
        \begin{center}
        \begin{tabular}{|c|c|c|}
        \hline
        \textbf{Algorithm} & \textbf{W} & \textbf{p-value} \\
        \hline
        Bubblesort & 0.934980 & \textbf{0.0666} \\
        \hline
        Insertion Sort & 0.949881 & \textbf{0.1678} \\
        \hline
        Mergesort & 0.936667 & \textbf{0.0739} \\
        \hline
        Quicksort & 0.936049 & \textbf{0.0712} \\
        \hline
        \end{tabular}
        \label{table-shapiro-test-ueq-001-100}
        \end{center}
    }
\end{table}

@@

\begin{figure}[H]
    \centering
    \frame{\includegraphics[scale=0.30]{figures/anova_prob_0_01_tam_100_Perc_largest_subarray_size.png}}
    \textsf{\caption[Q-Q plot for \%LSS with \textit{probability of failure} of 1\% and \textit{array size} of 100.]{Q-Q plot for \%LSS with \textit{probability of failure} of 1\% and \textit{array size} of 100.\label{fig-qqplot-lss-001-100}}}
 \end{figure}

@@

Tables \ref{table-shapiro-test-lss-005-1000} and \ref{table-shapiro-test-ueq-005-1000} shows another examples of results we got running Shapiro-Wilk test. The independent variables assume these values: \textit{probability of failure} of 0.05 and \textit{array size} of 1000. The bold \textit{p-values} confirms that the data is normally distributed.

\begin{table}[H]
    \parbox{.45\linewidth}{
        \caption{Shapiro-Wilk test for \%LSS with \textit{probability of failure} of 5\% and \textit{array size} of 1000.}
        \begin{center}
        \begin{tabular}{|c|c|c|}
        \hline
        \textbf{Algorithm} & \textbf{W} & \textbf{p-value} \\
        \hline
        Bubblesort & 0.891642 & 0.00052 \\
        \hline
        Insertion Sort & 0.921918 & 0.3000 \\
        \hline
        Mergesort & 0.878308 & 0.0025 \\
        \hline
        Quicksort & 0.917583 & 0.0232 \\
        \hline
        \end{tabular}
        \label{table-shapiro-test-lss-005-1000}
        \end{center}
    }
    \hfill
    \parbox{.45\linewidth}{
        \caption{Shapiro-Wilk test for \%UEQ with \textit{probability of failure} of 5\% and \textit{array size} of 1000.}
        \begin{center}
        \begin{tabular}{|c|c|c|}
        \hline
        \textbf{Algorithm} & \textbf{W} & \textbf{p-value} \\
        \hline
        Bubblesort & 0.936459 & \textbf{0.0730} \\
        \hline
        Insertion Sort & 0.931681 & \textbf{0.0544} \\
        \hline
        Mergesort & 0.961920 & \textbf{0.3465} \\
        \hline
        Quicksort & 0.963954 & \textbf{0.3892} \\
        \hline
        \end{tabular}
        \label{table-shapiro-test-ueq-005-1000}
        \end{center}
    }
\end{table}

Other examples of distribution normality can be seen in Q-Q plot in Figures \ref{fig-qqplot-lss-005-1000} and Y below. The considered dependent variable is \textit{percentage of the largest subarray size} with a \textit{probability of failure} value of 5\% and \textit{array size} value of 10000 for all sorting algorithms.

\begin{figure}[H]
    \centering
    \frame{\includegraphics[scale=0.30]{figures/anova_prob_0_05_tam_1000_Perc_largest_subarray_size.png}}
    \textsf{\caption[Q-Q plot for \%LSS with \textit{probability of failure} of 5\% and \textit{array size} of 1000.]{Q-Q plot for \%LSS with \textit{probability of failure} of 5\% and \textit{array size} of 1000.\label{fig-qqplot-lss-005-1000}}}
 \end{figure}

 \begin{figure}[H]
    \centering
    \frame{\includegraphics[scale=0.30]{figures/anova_prob_0_05_tam_1000_Perc_unordered_elements_quantity.png}}
    \textsf{\caption[Q-Q plot for \%UEQ with \textit{probability of failure} of 5\% and \textit{array size} of 1000.]{Q-Q plot for \%UEQ with \textit{probability of failure} of 5\% and \textit{array size} of 1000.\label{fig-qqplot-ueq-005-1000}}}
 \end{figure}

 \subsection{Hypothesis Testing}

 For hypothesis testing, we define a confidence degree of 95\%. As mentioned before, we only tested the hypothesis related to the dependent variable \%UEQ \textit{percentage of unordered elements quantity}. The hypothesis was:
 \begin{itemize}
     \item \textbf{Hypothesis 1}: For a given probability of failure and array size, tested algorithms will produce a different percentage of unordered elements quantity;
     \item \textbf{Hypothesis 3}: For each algorithm, the array size and probability of failure have a significative impact on the percentage of unordered elements quantity.
 \end{itemize}

 Because our dataset has 3 independent variables, we perform ANOVA method to simplify the hypothesis testing. This method allows comparing various groups of variables at the same time.

 \subsubsection{Testing Hypothesis 1}

We formulated the following hypothesis for the combination of independent variables \textit{probability of failure} and \textit{array size}:

\setcounter{hyp}{-1}
\begin{hyp} \label{hyp:a} There is no significant difference between algorithms related to variable \%UEQ.\end{hyp}
\begin{hyp} \label{hyp:b} There is significant difference between algorithms related to variable \%UEQ.\end{hyp}

After we run the testing method, we got the following results (Table \ref{tab-hypothesis1-testing}):

\begin{table}[H]
    \caption{Hypothesis 1 ANOVA results.}
    \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 1\% and Array Size = 100}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 1174.466667 & 3.0 & 171.368721 & \textbf{1.845762e-42} \\
    \hline
    residual & 265.000000 & 116.0 & & \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 1\% and Array Size = 1000}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 1635.180250 & 3.0 & 1817.615579 & \textbf{2.610034e-97} \\
    \hline
    residual & 34.785667 & 116.0 & & \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 1\% and Array Size = 10000}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 1948.204989 & 3.0 & 15831.708335 & \textbf{2.332995e-151} \\
    \hline
    residual & 4.758210 & 116.0 & & \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 2\% and Array Size = 100}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 1106.5 & 3.0 & 88.617785 & \textbf{7.054900e-30} \\
    \hline
    residual & 482.8 & 116.0 & & \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 2\% and Array Size = 1000}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 2286.83025 & 3.0 & 1990.144336 & \textbf{1.507899e-99} \\
    \hline
    residual & 44.43100 & 116.0 & & \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 2\% and Array Size = 10000}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 3081.306943 & 3.0 & 18442.105239 & \textbf{3.406264e-155} \\
    \hline
    residual & 6.460427 & 116.0 & & \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 5\% and Array Size = 100}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 2274.291667 & 3.0 & 225.58173 & \textbf{3.104916e-48} \\
    \hline
    residual & 389.833333 & 116.0 & & \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 5\% and Array Size = 1000}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 3704.037583 & 3.0 & 1202.21591 & \textbf{3.625400e-87} \\
    \hline
    residual & 119.132333 & 116.0 & & \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Probablity of Failure = 5\% and Array Size = 10000}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    algorithm & 4948.821397 & 3.0 & 19487.120298 & \textbf{1.402080e-156} \\
    \hline
    residual & 9.819533 & 116.0 & & \\
    \hline
    \end{tabular}
    \label{tab-hypothesis1-testing}
    \end{center}
\end{table}

Based on these results, we reject the null hypothesis ($H_{0}$) for all testing combinations. Therefore, we can affirm with 95\% of confidence that there is a significant difference between algorithms related do variable \%UEQ.

\subsubsection{Testing Hypothesis 3}

We formulated the following hypothesis for all algorithms considered in this work:

\setcounter{hyp}{-1}
\begin{hyp} \label{hyp:a} The values of variables \textit{probability of failure} and \textit{array size} don't have a significant impact on values of \%UEQ variable.\end{hyp}
\begin{hyp} \label{hyp:b} The values of variables \textit{probability of failure} and \textit{array size} have a significant impact on values of \%UEQ variable.\end{hyp}

After we run the testing method, we got the following results (Table \ref{tab-hypothesis3-testing}):

\begin{table}[H]
    \caption{Hypothesis 3 ANOVA results}
    \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Bubblesort}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    size\_of\_array & 0.019548 & 1.0 & 0.01350 & \textbf{0.90759} \\
    \hline
    probability\_of\_failure & 3590.644437 & 1.0 & 2479.71637 & \textbf{7.570027e-137} \\
    \hline
    residual & 385.169623 & 266.0 & &  \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Mergesort}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    size\_of\_array & 1784.873256 & 1.0 & 383.585800 & \textbf{1.702008e-53} \\
    \hline
    probability\_of\_failure & 13009.963175 & 1.0 & 2795.961630 & \textbf{3.801570e-143} \\
    \hline
    residual & 1237.731651 & 266.0 & &  \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Quicksort}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    size\_of\_array & 121.551362 & 1.0 & 32.964196 & \textbf{2.556033e-08} \\
    \hline
    probability\_of\_failure & 5626.955128 & 1.0 & 1526.005558 & \textbf{3.457996e-112} \\
    \hline
    residual & 980.841817 & 266.0 & &  \\
    \hline
    \multirow{2}{*}{\hfill}&\multicolumn{4}{|c|}{\textbf{Insertion sort}} \\
    \cline{2-5}
    &\textbf{\textit{sum\_sq}} & \textbf{\textit{df}} & \textbf{\textit{F}} & \textbf{\textit{PR(>F)}} \\
    \hline
    size\_of\_array & 309.843991 & 1.0 & 75.905868 & \textbf{3.236586e-16} \\
    \hline
    probability\_of\_failure & 6806.075083 & 1.0 & 1667.358570 & \textbf{1.414758e-116} \\
    \hline
    residual & 1085.798822 & 116.0 & & \\
    \hline
    \end{tabular}
    \label{tab-hypothesis3-testing}
    \end{center}
\end{table}

Based on these results, we reject the null hypothesis ($H_{0}$) for algorithms Mergesort, Quicksort, and Insertion sort. Therefore, we can affirm with 95\% of confidence that the values of variables \textit{probability of failure} and \textit{array size} have a significant impact on values of \%UEQ variable.

For Bubblesort algorithm, however, there is no confidence to reject the null hypothesis ($H_{0}$). Analyzing the result, we can affirm with 95\% of confidence that the value of \textit{probability of failure} variable had a significant impact on the value of the dependent variable \%UEQ. But we can't affirm the same related to independent variable \textit{array size}. These results demonstrate that, for Bubblesort algorithm, the array sizes used in this experiment were not decisive to define values concerning the percentage of unordered elements quantity.

\subsection{Performance analysis related to \textit{percentage of unordered elements quantity} variable}

In this section, we present an analysis of overall performance related to dependent variable \textit{percentage of unordered elements quantity}. To help this analysis, we show below 3 bar graphs, each of these for a different value of \textit{probability of failure} independent variable. These graphs shows informations about values of \%UEQ variable grouped by \textit{array size} and \textit{sorting algorithm} variables. Figures \ref{fig-barplot-pof-001}, \ref{fig-barplot-pof-002}, and \ref{fig-barplot-pof-005} below exhibit these graphs.

The bars in each graph represents the sample mean value for variable \%UEQ inside each group. Above each bar, there is a vertical line that indicates the lower and upper limits of a confidence interval for a significance level of 5\% ($\alpha = 0.05$).

\begin{figure}[H]
    \centering
    \frame{\includegraphics[scale=1.15]{figures/Barplot_Estatisticas_por_Algoritmo_Prob_0_01.png}}
    \textsf{\caption[Data for \textit{probability of failure} of 1\%.]{Data for \textit{probability of failure} of 1\%.\label{fig-barplot-pof-001}}}
 \end{figure}

 \begin{figure}[H]
    \centering
    \frame{\includegraphics[scale=1.15]{figures/Barplot_Estatisticas_por_Algoritmo_Prob_0_02.png}}
    \textsf{\caption[Data for \textit{probability of failure} of 2\%.]{Data for \textit{probability of failure} of 2\%.\label{fig-barplot-pof-002}}}
 \end{figure}

 \begin{figure}[H]
    \centering
    \frame{\includegraphics[scale=1.15]{figures/Barplot_Estatisticas_por_Algoritmo_Prob_0_05.png}}
    \textsf{\caption[Data for \textit{probability of failure} of 5\%.]{Data for \textit{probability of failure} of 5\%.\label{fig-barplot-pof-005}}}
\end{figure}

Considering the data in the graphs, we can conclude that Bubblesort generates lower mean value for variable \%UEQ for all defined combinations (\textit{probability of failure} X \textit{array size} X \textit{sorting algorithm}). These data demonstrate that Bubblesort was the less affected algorithm by the memory faults simulated in this experiment. We believe that this fact happens because of the operation of the algorithm itself, where all elements order are verified each iteration, causing an element incorrectly positioned in an iteration (because of a memory fault) can have its location corrected in an upcoming iteration.

Related to Bubblesort yet, we could identify that despite different sizes of input arrays (100, 1000, and 10000), there wasn't a significant impact on the mean value obtained for \%UEQ variable, proving what we verify with ANOVA when doing hypothesis 3 testing for this algorithm.

### 2019-12-22 paper/ExperimentalSetup.tex
finalização da escrita do artigo

We first state our problem, then describe our data, showing all its characteristics. Then, we provide our hypothesis, define the setup, and perform the testing. Finally, we collect and perform data analysis.

@@

% \setcounter{hyp}{-1}
% \begin{hyp}[Test hypothesis] \label{hyp:a}There is no difference in memory faults between tested algorithms.\end{hyp}
% \begin{hyp} \label{hyp:b}There are differences in memory faults between tested algorithms. \end{hyp}
% \begin{hyp} \label{hyp:c} An algorithm is better than others considering all dependent variables. \end{hyp}

% \begin{hyp} \label{hyp:d} ... \end{hyp}

@@

% \begin{itemize}
%     \item A script that generates input files;
%     \item Four files, one for each of those algorithms: quicksort, bubblesort, insertion sort, and mergesort, that are used to sorting the input data;
% \end{itemize}

@@

% We run 1000 times each algorithm with each file listed in Table \ref{table-input-data} below. We chose this number of executions to reduce possible noise in the generated data. Table \ref{table-input-data} shows 9 different input files with its sequence sizes and probabilities of failure. Plus, the number of executions and each sorting algorithm.

% For example, Input A showed in that table with its characteristics was executed 1000 times for each algorithm -  bubblesort, quicksort, mergesort, and insertion sort - totalizing 4000 executions.

% \begin{table}[H]
%     \caption{Generated input data.}
%     \begin{center}
%     \begin{tabular}{|c|c|c|c|c|}
%     \hline
%     \textbf{Input} & \textbf{Sequence Size} & \textbf{Probability of Failure} & \textbf{Executions} & \textbf{Algorithm} \\ 
%     \hline
%     \multirow{4}{*}{Input A} & \multirow{4}{*}{100} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input B} & \multirow{4}{*}{100} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input C} & \multirow{4}{*}{100} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input D} & \multirow{4}{*}{1000} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input E} & \multirow{4}{*}{1000} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input F} & \multirow{4}{*}{1000} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input G} & \multirow{4}{*}{10000} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input H} & \multirow{4}{*}{10000} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input I} & \multirow{4}{*}{10000} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \end{tabular}
%     \label{table-input-data}
%     \end{center}
% \end{table}

% Each execution generates an output file. Then, we generate 4000 outputs for each combination of sequence size, and the probability of failure, totalizing 36000 files. Figure \ref{fig-output-file-example} shows an output file.

@@

After generating the dataset, we developed a Python script that reads the 1080 output files and produces a single CSV file (first lines showed in Figure \ref{fig-output-csv-file-example} below), containing the following columns:
\begin{itemize}
    \item \textit{algorithm}: the algorithm used to sort the array;
    \item \textit{probability\_of\_failure}: the probability of failure used when sorting;
    \item \textit{size\_of\_array}: the size of the array to be sorted;
    \item \textit{largest\_sorted\_subarray}: the largest sorted subarray after sorting;
    \item \textit{k\_unordered\_sequence}: number of unordered sequence after sorting.
    \item \textit{percentage\_k\_unordered}: percentage of unordered sequence after sorting related to original array;
    \item \textit{percentage\_largest\_sorted\_subarray}: percentage of largest sorted subarray after sorting related to original array.
\end{itemize}

@@

% \subsection{Data Analysis}

% In this section, we describe our data and the statistic tests used to test the hypothesis showed before.

% \subsubsection{Exploratory Data Analysis (EDA)}

% Based on the output files generated before, our starting point was to test if the dataset had a normal distribution. To achieve this, we use Shapiro-Wilk normality test. After we run these tests, only the distribution related to the dependent variable \textit{unordered elements size} was considered normal. Figure \ref{fig-distribution-graph} shows an example of this distribution. In these graphs, the red vertical line means the mean, and the blue vertical line means the median. The distributions were normal in all combinations of independent variables for dependent variable \textit{unordered elements size}.

% \begin{figure}[H]
%     \centering
%      \begin{subfigure}{.5\textwidth}
%      \centering
%      \frame{\includegraphics[scale=0.12]{figures/0_01_10000_Bubble_distribution_graph}}
%      \textsf{\caption[Distribution Graph for bubblesort.]{Distribution Graph for bubblesort.\label{fig-dist-bubble-001-100}}}
%      \end{subfigure}%
%      \begin{subfigure}{.5\textwidth}
%      \centering
%      \frame{\includegraphics[scale=0.12]{figures/0_01_10000_Insertion_distribution_graph.png}}
%      \textsf{\caption[Distribution Graph for insertion sort.]{Distribution Graph for insertion sort.\label{fig-dist-insertion-001-100}}}
%      \end{subfigure}
%      \begin{subfigure}{.5\textwidth}
%      \centering
%      \frame{\includegraphics[scale=0.12]{figures/0_01_10000_Merge_distribution_graph.png}}
%      \textsf{\caption[Distribution Graph for mergesort.]{Distribution Graph for mergesort.\label{fig-dist-merge-001-100}}}
%      \end{subfigure}%
%      \begin{subfigure}{.5\textwidth}
%      \centering
%      \frame{\includegraphics[scale=0.12]{figures/0_01_10000_Quick_distribution_graph.png}}
%      \textsf{\caption[Distribution Graph for quicksort.]{Distribution Graph for quicksort.\label{fig-dist-quick-001-100}}}
%      \end{subfigure}
%      \caption{Distribution graph for a probability failure of 1\% and a sequence size of 10000.}
%     \label{fig-distribution-graph}
% \end{figure}

% Q-Q plot shows that how much more blue points close to the red line, most normal is the distribution. Figure \ref{fig-qqplot} below presents this graph for a probability failure of 1\% and a sequence size of 10000 for considered sorting algorithms.

% \begin{figure}[H]
%     \centering
%     \includegraphics[scale=0.37]{figures/anova_prob_0_01_tam_10000_DESORDENADOS_PERCENTUAL.png}
%     \textsf{\caption[Q-Q plot showing normal distribution for a probability failure of 1\% and a sequence size of 10000.]{Q-Q plot showing normal distribution for a probability failure of 1\% and a sequence size of 10000.\label{fig-qqplot}}}
%  \end{figure}

\subsection{Conclusions}

>>>>

In this section, we present the setup of this study. We first state our problem, then define the independent and dependent variables. After that, we show the hypothesis and describe our dataset, showing all its characteristics.

@@

@@

@@

@@

After generating the dataset, we developed a Python script that reads the 1080 output files and produces a single CSV file (first lines showed in Figure \ref{fig-output-csv-file-example} below). This file was used to run the statistical tests, and contains the following columns:
\begin{itemize}
    \item \textit{algorithm}: the algorithm used to sort the array;
    \item \textit{probability\_of\_failure}: the probability of failure used when sorting;
    \item \textit{size\_of\_array}: the size of the array to be sorted;
    \item \textit{largest\_sorted\_subarray}: the largest sorted subarray after sorting;
    \item \textit{k\_unordered\_sequence}: number of unordered sequence after sorting.
    \item \textit{percentage\_k\_unordered}: percentage of unordered sequence after sorting related to original array;
    \item \textit{percentage\_largest\_sorted\_subarray}: percentage of largest sorted subarray after sorting related to original array.
\end{itemize}

@@
